A simplified dynamic bioenergetic model for coral bleaching

### 2016-10-13 JTB_ms/JTB_ms.tex
update new C feedback

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A simplified dynamic bioenergetic model for coral-Symbiodinium symbioses},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{A simplified dynamic bioenergetic model for coral-\emph{Symbiodinium}
symbioses}
    \author[University of Hawaii]{Ross Cunning\corref{c1}}
   \ead{ross.cunning@gmail.com} 
   \cortext[c1]{Corresponding Author}
    \author[University of California]{Erik B. Muller}

@@

  \begin{abstract}
  This is the abstract.
  \end{abstract}

@@

@@

\(j_{CO_2}^p\)

@@

Passive CO\textsubscript{2} delivery to symbiont

@@

4.04e-3

@@

\(j_{CO_2}^a\)
Active CO\textsubscript{2} delivery to symbiont
0.32--18.0

@@

\(n_{LC}\)

@@

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig1.png}
\caption{Graphical representation of coral-algal symbiosis model. Light,
CO\textsubscript{2}, prey, and DIN are acquired from the external
environment proportional to the biomass of the partner indicated by the
black box for uptake. Mass fluxes (see Table 1 for definitions) are
represented by \(j\)'s with subscripts indicating the type of mass, and
in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
shared by one partner with the other. Parallel complementary
synthesizing units (SUs) are represented by large circles, and rejection
fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
proportional rate that impacts other model fluxes by inhibition or
acceleration. Recycled mass fluxes from biomass turnover are not shown
for clarity (but see Table 1 for definitions).}
\end{figure}

@@

Due to the inherent inefficiency of the parallel complementary SU
formulation, there is always some nitrogen shared with the symbiont even
when coral biomass formation is strongly nitrogen-limited. Likewise,
there is always a non-zero rejection flux of carbon from the coral
biomass SU, which is assumed to be lost to the environment.

@@

We then specify two pathways for input of inorganic carbon to the
photosynthesis SU: 1) passive diffusion of CO\textsubscript{2} from the
external environment (\(j_{CO_2}^p\)), and 2) active delivery of
CO\textsubscript{2} to the symbiont by the host (\(j_{CO_2}^a\)). The
passive flux ensures that some CO\textsubscript{2} is always available
to photosynthesis, and the active flux encompasses potentially diverse
mechanisms by which the host may enhance CO\textsubscript{2}
availability for the symbiont, including active transport of
bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
CO\textsubscript{2} to promote diffusion toward the symbiont (Tansik,
Fitt, and Hopkinson 2015), and acidification of the symbiosome to
increase localized CO\textsubscript{2} concentrations around the
symbiont (Barott et al. 2014). Since the host physically separates the
symbiont from the external environment, both the passive and active flux
rates are proportional to host biomass. A more rigorous, mechanistic
model of inorganic carbon processing would require spatially explicit
internal pools accounting for pH and carbon speciation, which is beyond
the current scope of this work. Instead, the specification of fixed
passive and active CO\textsubscript{2} input rates, \(j_{CO_2}^p\) and
\(j_{CO_2}^a\), offers the user the opportunity to compare different
rates of CO\textsubscript{2} delivery that may characterize different
coral species (Wooldridge 2014a). The input of CO\textsubscript{2} to
the photosynthesis SU is therefore specified as:

\begin{equation} j_{CO_2} = j_{CO_2}^p + j_{CO_2}^a \end{equation}

@@

\begin{equation} j_{eL} = (j_L - j_{CP} / n_{LC})_+ \end{equation}

This excess light energy must be quenched by alternative pathways in
order to prevent photooxidative damage (Powles 1984).
\emph{Symbiodinium} utilize a variety of pathways for non-photochemical
quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
specified as a parameter of the symbiont (\(k_{NPQ}\)). NPQ capacity may
also be susceptible to reduction by photoinhibition, so the total NPQ
capacity is specified as:

\begin{equation} j_{NPQ}= k_{NPQ} \cdot c_{ROS}^{-1} \end{equation}

@@

\section{Sensitivity analysis}\label{sensitivity-analysis}

\section{Steady state behavior}\label{steady-state-behavior}

To analyze qualitative behavior, we ran the model to steady state across
gradients of external irradiance and nutrients (Fig. 2), which revealed
patterns consistent with observed phenomena in corals. Predicted growth
rates are low at low light and DIN, and begin increasing as both of
these factors increase (Fig. 2A). Low light is limiting to
photosynthesis (Fig. 2D), which leads to carbon-limitation of host
growth (Fig. 2B) and an associated increase in the symbiont to host
biomass ratio (Fig. 2C). In agreement with this trend are many
observations of negative correlation between irradiance and symbiont
density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000; Titlyanov et
al. 2001). As higher light alleviates light-limitation of
photosynthesis, growth becomes less carbon-limited. Similarly, elevating
DIN up to \textasciitilde{}2 µM alleviates nitrogen-limitation (Fig.
2B). Increased growth at higher DIN is predicted by the DEB model of
Muller et al. (2009), and has also been observed experimentally
(Muller-Parker et al. 1994; Tanaka et al. 2007; Tanaka et al. 2013).
However, DIN elevation beyond a certain point (e.g., 2-3 µM in these
simulations) has little effect on growth as carbon becomes limiting
(Fig. 2B). Very high nutrient levels may even reduce growth (Shantz,
Lemoine, and Burkepile 2015), although these impacts are not likely to
occur within the range of concentrations considered here (\textless{}4
µM) (Ferrier-Pagès et al. 2000).

@@

\section{Dynamic behavior}\label{dynamic-behavior}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A simplified dynamic bioenergetic model for coral bleaching},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{A simplified dynamic bioenergetic model for coral bleaching}
    \author[University of Hawaii]{Ross Cunning\corref{c1}}
   \ead{ross.cunning@gmail.com} 
   \cortext[c1]{Corresponding Author}
    \author[University of California]{Erik B. Muller}

@@

  \begin{abstract}
  This is the abstract. Some type of positive feedback is necessary.
  Compare scenarios of recycling CO2, have photodamage increase
  maintenance that does recycle CO2. Compare recycled N to symbiont
  instead of host How much do you have to reduce the light to get
  recovery? i.e., magnitude of hysteresis?
  \end{abstract}

@@

\(j_{eC}\)
Excess carbon used to activate host CCMs

@@

\(k_{CO_2}\)

@@

Efficacy of CO\textsubscript{2} delivery by host CCMs

@@

@@

molCO\textsubscript{2} molC\textsuperscript{-1}

@@

\(y_{CL}\)

@@

\begin{figure}[htbp]
\centering
\includegraphics{../img/Fig1.png}
\caption{Graphical representation of coral-algal symbiosis model. Light,
CO\textsubscript{2}, prey, and DIN are acquired from the external
environment proportional to the biomass of the partner indicated by the
black box for uptake. Mass fluxes (see Table 1 for definitions) are
represented by \(j\)'s with subscripts indicating the type of mass, and
in some cases the process (e.g., \(j_{CP}\) is the flux of carbon
produced by photosynthesis), and \(\rho\)'s indicate fluxes that are
shared by one partner with the other. Parallel complementary
synthesizing units (SUs) are represented by large circles, and rejection
fluxes from these SUs are indicated by dashed lines. \(c_{ROS}\) is a
proportional rate that impacts other model fluxes by inhibition or
acceleration; likewise, \(j_{eC}\) accelerates the rate of \(j_{CO_2}\).
Recycled mass fluxes from biomass turnover are not shown for clarity
(but see Table 1 for definitions).}
\end{figure}

@@

The carbon rejected from this SU reflects the amount of excess fixed
carbon available to the host that is not used in biomass formation:

\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}

This flux, \(j_{eC}\), is assumed to be available as a respiratory
substrate to support various energetically-demanding processes; of
particular importance is the host's active supply of CO\textsubscript{2}
for symbiont photosynthesis (Wooldridge 2013; Hopkinson, Tansik, and
Fitt 2015) (see below).

Due to the inherent inefficiency of the parallel complementary SU
formulation, there is always some nitrogen shared with the symbiont even
when coral biomass formation is strongly nitrogen-limited. Likewise,
there is always a non-zero rejection flux of excess carbon from the
coral biomass SU.

@@

We then specify the input of CO\textsubscript{2} to photosynthesis as
\(j_{CO_2}\), which is mediated by the host. This encompasses
potentially diverse mechanisms by which the host may enhance
CO\textsubscript{2} availability for photosynthesis (i.e.,
carbon-concentrating mechanisms (CCMs)), including active transport of
bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to
CO\textsubscript{2} to promote diffusion toward the symbiont (Tansik,
Fitt, and Hopkinson 2015), and acidification of the symbiosome to
increase localized CO\textsubscript{2} concentrations around the
symbiont (Barott et al. 2014). Since these mechanisms require energetic
input by the host, the magnitude of CO\textsubscript{2} delivery is set
proportional to \(j_{eC}\), the amount of excess carbon from the host
biomass SU, assuming that some of this carbon is respired to energize
these CCMs. This formulation means that the symbiont indirectly ensures
its own CO\textsubscript{2} supply by providing the host with fixed
carbon (Wooldridge 2013), which establishes a positive feedback in the
model that proves to be essential in simulating the phenomenon of coral
bleaching. The parameter \(k_{CO_2}\) is included to scale the efficacy
of host CCMs, i.e., how many moles of CO\textsubscript{2} can be
delivered from the environment to the symbiont per mole of fixed carbon.
Adjusting this parameter offers the user the opportunity to compare
different rates of CO\textsubscript{2} delivery that may characterize
different coral species (Wooldridge 2014a). The input of
CO\textsubscript{2} to the photosynthesis SU is therefore specified as:

\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}

@@

\begin{equation} j_{eL} = (j_L - j_{CP} / y_{CL})_+ \end{equation}

This excess light energy must be quenched by alternative pathways in
order to prevent photooxidative damage (Powles 1984).
\emph{Symbiodinium} utilize a variety of pathways for non-photochemical
quenching (NPQ; Roth 2014), which we collect in a relative NPQ capacity
specified as a parameter of the symbiont (\(k_{NPQ}\)). The total NPQ
flux is specified as a single-substrate SU formula:

\begin{equation} j_{NPQ} = \bigg({1 \over k_{NPQ}} + {1 \over j_{eL}}\bigg)^{-1} \end{equation}

@@

\section{Results}\label{results}

\subsection{Qualitative dynamics}\label{qualitative-dynamics}

simulations in constant environs show either positive or negative
exponential growth. some environments show two steady states, one with
positive and one with negative growth, depending on initial conditions.
based on numerical analysis, we see ultimate exponential growth or
decline. the mechanism that produces these alternate steady states is:
\ldots{}. . for practical purposes, we are interested in the positive
growth steady states and their properties and how they change across
environmental gradients (see section X). but we are also interested in
the possibility for flips to the alternate steady states (see section
Y).

\subsection{Steady state behavior}\label{steady-state-behavior}

\emph{emphasize that steady state is exponential growth, and fluxes are
per unit biomass/specific rates} To analyze qualitative behavior, we ran
the model to steady state across gradients of external irradiance and
nutrients (Fig. 2), which revealed patterns consistent with observed
phenomena in corals. Predicted growth rates are low at low light and
DIN, and begin increasing as both of these factors increase (Fig. 2A).
Low light is limiting to photosynthesis (Fig. 2D), which leads to
carbon-limitation of host growth (Fig. 2B) and an associated increase in
the symbiont to host biomass ratio (Fig. 2C). In agreement with this
trend are many observations of negative correlation between irradiance
and symbiont density (Stimson 1997; Brown et al. 1999; Fitt et al. 2000;
Titlyanov et al. 2001). As higher light alleviates light-limitation of
photosynthesis, growth becomes less carbon-limited. Similarly, elevating
DIN up to \textasciitilde{}2 µM alleviates nitrogen-limitation (Fig.
2B). Increased growth at higher DIN is predicted by the DEB model of
Muller et al. (2009), and has also been observed experimentally
(Muller-Parker et al. 1994; Tanaka et al. 2007; Tanaka et al. 2013).
However, DIN elevation beyond a certain point (e.g., 2-3 µM in these
simulations) has little effect on growth as carbon becomes limiting
(Fig. 2B). Very high nutrient levels may even reduce growth (Shantz,
Lemoine, and Burkepile 2015), although these impacts are not likely to
occur within the range of concentrations considered here (\textless{}4
µM) (Ferrier-Pagès et al. 2000).

@@

\subsection{Sensitivity analysis -- should come after the steady states
bc describes changes in steady
states}\label{sensitivity-analysis-should-come-after-the-steady-states-bc-describes-changes-in-steady-states}

\subsection{Dynamic behavior}\label{dynamic-behavior}
